--
title: "Reproducible Research Peer Assessment 1"
author: "Usha"
date: "2024-06-09"
output: html_document
---

# Loading and preprocessing the data
```{r load-data, echo=TRUE}
library(ggplot2)
library(dplyr)

# Load the data
data <- read.csv("activity.csv")

# Convert the date column to Date class
data$date <- as.Date(data$date, format="%Y-%m-%d")

## What is mean total number of steps taken per day?
```{r total-steps-per-day, echo=TRUE}
total_steps_per_day <- data %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarise(total_steps = sum(steps))
```{r histogram-steps-per-day, echo=TRUE, fig.height=5, fig.width=7}
ggplot(total_steps_per_day, aes(x=total_steps)) +
    geom_histogram(binwidth=1000, fill="blue", color="black") +
    xlab("Total Steps per Day") +
    ylab("Frequency") +
    ggtitle("Histogram of Total Steps per Day")
```{r histogram-steps-per-day, echo=TRUE, fig.height=5, fig.width=7}
ggplot(total_steps_per_day, aes(x=total_steps)) +
    geom_histogram(binwidth=1000, fill="blue", color="black") +
    xlab("Total Steps per Day") +
    ylab("Frequency") +
    ggtitle("Histogram of Total Steps per Day")
```{r mean-median-steps-per-day, echo=TRUE}
mean_steps <- mean(total_steps_per_day$total_steps)
median_steps <- median(total_steps_per_day$total_steps)

mean_steps
median_steps

## What is the average daily activity pattern?
```{r average-steps-interval, echo=TRUE}
average_steps_interval <- data %>%
    filter(!is.na(steps)) %>%
    group_by(interval) %>%
    summarise(average_steps = mean(steps))
```{r time-series-plot, echo=TRUE, fig.height=5, fig.width=7}
ggplot(average_steps_interval, aes(x=interval, y=average_steps)) +
    geom_line() +
    xlab("5-minute Interval") +
    ylab("Average Steps") +
    ggtitle("Average Daily Activity Pattern")
```{r max-interval, echo=TRUE}
max_interval <- average_steps_interval[which.max(average_steps_interval$average_steps),]
max_interval
## Imputing missing values
```{r total-na, echo=TRUE}
total_na <- sum(is.na(data$steps))
total_na
```{r fill-na, echo=TRUE}
data_filled <- data
for (i in 1:nrow(data_filled)) {
    if (is.na(data_filled$steps[i])) {
        data_filled$steps[i] <- average_steps_interval$average_steps[which(average_steps_interval$interval == data_filled$interval[i])]
    }
}
```{r total-steps-per-day-filled, echo=TRUE, fig.height=5, fig.width=7}
total_steps_per_day_filled <- data_filled %>%
    group_by(date) %>%
    summarise(total_steps = sum(steps))

ggplot(total_steps_per_day_filled, aes(x=total_steps)) +
    geom_histogram(binwidth=1000, fill="blue", color="black") +
    xlab("Total Steps per Day") +
    ylab("Frequency") +
    ggtitle("Histogram of Total Steps per Day (Filled Data)")
```{r mean-median-steps-per-day-filled, echo=TRUE}
mean_steps_filled <- mean(total_steps_per_day_filled$total_steps)
median_steps_filled <- median(total_steps_per_day_filled$total_steps)

mean_steps_filled
median_steps_filled
## Are there differences in activity patterns between weekdays and weekends?
```{r weekday-weekend, echo=TRUE}
data_filled$day_type <- ifelse(weekdays(data_filled$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
data_filled$day_type <- as.factor(data_filled$day_type)
```{r average-steps-day-type, echo=TRUE}
average_steps_day_type <- data_filled %>%
    group_by(interval, day_type) %>%
    summarise(average_steps = mean(steps))
```{r panel-plot, echo=TRUE, fig.height=7, fig.width=7}
ggplot(average_steps_day_type, aes(x=interval, y=average_steps, color=day_type)) +
    geom_line() +
    facet_wrap(~day_type, ncol=1) +
    xlab("5-minute Interval") +
    ylab("Average Steps") +
    ggtitle("Average Daily Activity Pattern by Day Type")

## Conclusion

This concludes the creation of the R Markdown file.
---

---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
